### 2025-01-31 paper/InBrief.qmd
Initial commit for submission

---
title: "msqrob2TMT: robust linear mixed models for inferring differential abundant proteins in labelled experiments with arbitrarily complex design - In Brief"
authors:
  - name: Stijn Vandenbulcke
    affiliations:
      - ref: VIB
      - ref: Biomed
      - ref: UGent
  - name: Christophe Vanderaa
    affiliations:
      - ref: UGent
  - name: Oliver Crook
    affiliations:
      - ref: UK
  - name: Lennart Martens
    affiliations:
      - ref: VIB
      - ref: Biomed
  - name: Lieven Clement
    affiliations:
      - ref: UGent
    corresponding: true

affiliations:
  - id: VIB
    name: VIB-UGent Center for Medical Biotechnology, VIB, Ghent, Belgium
  - id: Biomed
    name: Department of Biomolecular Medicine, Ghent University, Ghent, Belgium
  - id: UGent
    name: Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent, Belgium
  - id: UK
    name: Department of Statistics, University of Oxford, Oxford, UK.

bibliography: references.bib

format:
  pdf:
    mathspec: true
    number-sections: true
    number-depth: 3
    pdf-engine: latexmk
    pdf-engine-opt: -xelatex
    keep-tex: true
    include-in-header:
      text:
        \usepackage{amsmath, booktabs, caption, longtable, xcolor, todonotes}
editor: source
crossref:
  fig-title: Figure     # (default is "Figure")
  fig-prefix: Figure
---

# In Brief

Novel msqrob2TMT workflows for differential abundance analysis of labelled proteomics data with simple and complex designs. Incorporating modular preprocessing and flexible modelling using multiple continuous and categorical variables, and their interactions. Accounting for correlations within spectra, channels, runs, mixtures, and in longitudinal designs using random effects.
Improved performance of our peptide-level and protein-level robust regression workflows over existing methods, demonstrated in two spike-in studies and one case study. Accessible through open-source Bioconductor package msqrob2 and companion GitHub page: https://github.com/statOmics/msqrob2TMT_paper.