### 2025-01-31 paper/highlights.qmd

---
title: "msqrob2TMT: robust linear mixed models for inferring differential abundant proteins in labelled experiments with arbitrarily complex design - Highlights"
authors:
  - name: Stijn Vandenbulcke
    affiliations:
      - ref: VIB
      - ref: Biomed
      - ref: UGent
  - name: Christophe Vanderaa
    affiliations:
      - ref: UGent
  - name: Oliver Crook
    affiliations:
      - ref: UK
  - name: Lennart Martens
    affiliations:
      - ref: VIB
      - ref: Biomed
  - name: Lieven Clement
    affiliations:
      - ref: UGent
    corresponding: true

affiliations:
  - id: VIB
    name: VIB-UGent Center for Medical Biotechnology, VIB, Ghent, Belgium
  - id: Biomed
    name: Department of Biomolecular Medicine, Ghent University, Ghent, Belgium
  - id: UGent
    name: Department of Applied Mathematics, Computer Science and Statistics, Ghent University, Ghent, Belgium
  - id: UK
    name: Department of Statistics, University of Oxford, Oxford, UK.

bibliography: references.bib

format:
  pdf:
    mathspec: true
    number-sections: true
    number-depth: 3
    pdf-engine: latexmk
    pdf-engine-opt: -xelatex
    keep-tex: true
    include-in-header:
      text:
        \usepackage{amsmath, booktabs, caption, longtable, xcolor, todonotes}
editor: source
crossref:
  fig-title: Figure     # (default is "Figure")
  fig-prefix: Figure
---

# Highlights

- Statistical workflows for differential abundance analysis in labelled proteomics.
- Enhanced sensitivity and specificity.
- Flexible modelling for experiments with both simple and complex designs.
- Benchmarked against DEqMS, MSstatsTMT, and msTrawler using two spike-in datasets.
- Transparent and modular workflows through open-source Bioconductor package msqrob2.